Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY

### 2022-02-11 manuscript/draft-manuscript.Rmd
include ineligible treated people in match summary

## Study population

A total of `r label_number(1, big.mark=",")(filter(flowchart, crit=="c0") %>% pull(n))` fully vaccinated (received a primary course of two COVID-19 vaccine doses) people aged 18 and over and registered at a TPP practice on `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` were identified the database, with `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` (`r label_percent(0.1)(filter(flowchart, crit=="c4") %>% pull(pct_all))`) eligible for inclusion (supplementary Tables S1 and S2). Of these, `r label_number(1, big.mark=",")(cohortsummary$n_boosted)` received a booster dose during the recruitment period (up to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`). `r label_number(1, big.mark=",")(cohortsummary$n_3pfizer)` ( `r label_percent(0.1)(cohortsummary$prop_3pfizer)`) for BNT162b2 and `r label_number(1, big.mark=",")(cohortsummary$n_3moderna)` ( `r label_percent(0.1)(cohortsummary$prop_3moderna)`) for mRNA-1273 (Figure 1).

## Matching

`r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_recruited))`) BNT162b2 booster recipients were successfully matched to unboosted controls, contributing `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms. For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).

@@

Booster vaccination update over the duration of the study cohort, and the proportion of those that were matched for modelling.

@@

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
matchcoverage  %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2x BNT162b2  / 1x BNT162b2",
      vax12_type=="az-az" & treatment=="pfizer" ~ "2 x ChAdOx1-S / 1 x BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2x BNT162b2 / 1x mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2x ChAdOx1-S / 1x mRNA-1273",
      TRUE ~ NA_character_
    )
  ) %>%
  ggplot() +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=fct_rev(matched),
      fill=fct_rev(matched),
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.8,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL
cat("  \n\n")
```

>>>>

## Study population and matching

A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2). Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1), with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).

<!-- For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).  -->

@@

Booster vaccination uptake over the duration of the study period, and the proportion of those that were matched for modelling.

@@

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
matchcoverage  %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2x BNT162b2  / 1x BNT162b2",
      vax12_type=="az-az" & treatment=="pfizer" ~ "2 x ChAdOx1-S / 1 x BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2x BNT162b2 / 1x mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2x ChAdOx1-S / 1x mRNA-1273",
      TRUE ~ NA_character_
    )
  ) %>%
  ggplot() +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=status,
      fill=status,
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.8,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL
cat("  \n\n")
```
